---
title: "Mobula Tag maps"
author: "Jennifer Stahl"
date: "August 12, 2019"
output:
  pdf_document: default
  html_document: default
---
##Load packages
Choose the California cran mirror "51" to install packages. 
Install ggmap to download google maps, ggplot2 to plot google maps with points, lines, and polygons, etc., and tidyverse to glimpse data.
```{r message=FALSE, warning=FALSE}
chooseCRANmirror(ind=51) 
install.packages("ggmap") 
install.packages("ggplot2") 
install.packages("tidyverse") 
install.packages("devtools")
devtools::install_github("dkahle/ggmap")
#install.packages ("rgdal") #package to read shapefile.
install.packages ("sf") #package to read shapefile.
install.packages("ggsn")#package for north arrow.
install.packages("cowplot")#cowplot is an add-in to ggplot
install.packages("ggspatial")
```

```{r message=FALSE, warning=FALSE}
library(ggmap)
library(ggplot2)
library(ggspatial)
library(tidyverse)
#library(devtools)
#library(rgdal) 
library(sf)
#library(ggsn)
library(cowplot) #for insert map
```
## Get base maps
###Register google key 
```{r include=FALSE, warning=FALSE}
register_google(key = "") 
```
### Import map of Pacific centered on Hawaii.
```{r import Pacific map}
Pacgooglemap <- get_map(location = c(lon = -155, lat = 20),
                       source   = "google",
                       maptype  = "satellite", #can also do "terrain" for a simplier map.
                       zoom     = 3)# This is the zoom parameter, larger number zooms in more.
```
###Display google map using ggmap, which uses grid graphics.
```{r display Pacific map}
Pacbasemap<- ggmap(Pacgooglemap)
Pacbasemap
```
###Import maps centered in Hawaii.
```{r import AS and HI maps}
HIgooglemap <- get_map(location = c(lon = -160, lat = 20),
                       source   = "google",
                       maptype  = "satellite",
                       zoom     = 4)
```
###Display google maps using ggmap.
```{r display maps}
HIbasemap<- ggmap(HIgooglemap)
HIbasemap

```

```{r}
mobula_tags <-read.csv(file="C:/R_git_stuff/Mobula-tag-maps/Mobula_locations.csv", header=TRUE, sep=",")
#head(mobula_tags)
#glimpse(mobula_tags)
#dim(mobula_tags) #8x12
```

#Change Tag_ID to character so can use this field to color different tag paths on the map
```{r}
mobula_tags$Ptt <- as.character(mobula_tags$Ptt)
```

##Map tag data.
###First map tag releases and recoveries as point data and look for outliers.
```{r map tag points}
tag_rel <- HIbasemap + geom_point(data = mobula_tags, aes(Deploy_long, Deploy_lat), colour = "yellow")
tag_rel 

tag_rec <- HIbasemap + geom_point(data = mobula_tags, aes(Popoff_long, Popoff_lat), colour = "red")
tag_rec 
#Note some tags may not display. Tags with no popoff lat/long (mortalities or tags at large). If tags are in eastern hemisphere then will need to run below code to scale map from 0-360 degrees and then rerun this code. Also tags may not display if they are out of the map bounds.
```

###Ggplot maps on a 0-360 scale so does not change to positive 180 when crossing the antimeridian. 
### Converted tags across the antimeridian to a 0-360 longitude, so they will display on the map. 
```{r correct longitudes}
mobula_tags$Popoff_long[which(mobula_tags$Popoff_long>0)] <- -360 + (mobula_tags$Popoff_long[which(mobula_tags$Popoff_long>0)])
```

#Tag map by species
###Map tag paths from release to recovery with release locations. (use this if want to map tags still at large)
```{r map tagpaths}
tagpaths <- HIbasemap + geom_point(data = mobula_tags, aes(Deploy_long, Deploy_lat), colour = "yellow") + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Species), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data = mobula_tags)
tagpaths
ggsave("mobula_tag_map_by_species.jpg") #
```

###Map tag paths only
```{r}
tagpaths <- HIbasemap + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Species), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data = mobula_tags)
tagpaths
ggsave("mobula_tag_map_by_species_paths_only.jpg") 
```


##Add map elements
###In ggspatial you can add a scale bar and map arrow.Need to work on scale bar.
ggplot() + 
geom_sf(data=nc) + #I think this does something with coordinate system.
annotation_scale(location = "tl") +  #'tr' here stands for top right and is the placement on the figure of the scale bar. bl, tl etc. work as well. 
annotation_north_arrow(location = "br", which_north = "true") # this is the arrow, and you can pick your placement and style too I believe. 

```{r}
#adds north arrow. 
tag_map_with_elements <- tagpaths + annotation_north_arrow(location = "bl", which_north = "true")
tag_map_with_elements
ggsave("mobula_tag_map_July29_2024.jpg") 
#help(annotation_north_arrow)
```

#get error with scale bar code
```{r}
#tag_map_with_elements <- tagpaths + annotation_north_arrow(location = "tr", which_north = "true") + geom_spatial_point(aes(x, y), crs = 4326) +
  #annotation_scale(location = "tl") +
  #coord_sf(crs = 3995)

#tag_map_with_elements

#help(geom_sf)
#help(annotation_scale)
```

#Subset tag data by species and/or tag to get maps of tags with corresponding depth data.
```{r}
Manta <-mobula_tags %>% filter(Species == "M birostris")
Tarapacana <-mobula_tags %>% filter(Species == "M tarapacana")
#Mobular <-mobula_tags %>% filter(Species == "M mobular")
Mobular_234811 <-mobula_tags %>% filter(Tag_ID == 234811)
Thurstoni <-mobula_tags %>% filter(Species == "M thurstoni")
#Thurstoni <-mobula_tags %>% filter(Tag_PTT == 234821|Tag_PTT == 234820)
```

```{r}
Manta_path <- HIbasemap + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Species), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data = Manta)
ggsave("Manta_path.jpg") 
```

```{r}
Tarapacana_path <- HIbasemap + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Species), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data = Tarapacana)
ggsave("Tarapacana_path_path.jpg") 
```

```{r}
Mobular_234811_path <- HIbasemap + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Species), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data = Mobular_234811)
ggsave("Mobular_234811_path.jpg")

#This has release info for both tags and recovery for one that survived.
#Mobular_path <- HIbasemap + geom_point(data=Mobular, aes(Deploy_long, Deploy_lat), colour = "yellow") + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Tag_ID), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data=Mobular)
```

```{r}
Thurstoni_path <- HIbasemap + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Tag_ID), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data =Thurstoni)
ggsave("Thurstoni_path.jpg")

#This adds the release locations as points.
#Thurstoni_path <- HIbasemap + geom_point(data = Thurstoni, aes(Deploy_long, Deploy_lat), colour = "yellow") + geom_segment(aes(x = Deploy_long, y = Deploy_lat, xend = Popoff_long, yend = Popoff_lat, colour = Tag_ID), arrow = arrow(angle = 30, length = unit(0.2, "cm"), ends = "last", type = "open"), data =Thurstoni)

```

#Distance between 2 points. Was having trouble with this so used NOAA online calculator.
https://www.nhc.noaa.gov/gccalc.shtml
```{r}
#pointDistance(c(17.495,-163.0605), c(36.3941,176.9806), lonlat=TRUE) #can't figure out the package this goes to. 

library ("geosphere") 
#getting error when run this. 
#answer in meters.

#distm (c(17.495,-163.0605), c(36.3941,176.9806), fun = distGeo) #manta
#distm (c(28.94295,-156.089), c(22.5146,-157.712), fun = distGeo)#tarapacana
#distm (c(12.01667, -169.267), c(14.6846, -171.5294), fun = distGeo)#thurstoni, 22P0709, 234821
#distm (c(16.01667, -158.600), c(16.9736, -150.9331), fun = distGeo)#thurstoni, 22P0707, 234820
```

